---
title: "How to use flipR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use flipR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we'll discuss the functions and data contained within this package
and explain the use cases for each component. 

## Introduction 

This package was designed to aide users in manipulating the order and orientation 
of data frames. The namesake function of this package, `flip()`, was inspired by data
generated when analyzing model adequacy of a set of character data to a phylogenetic
tree. Included in this package is a data frame containing a snippet of the data
generated by said analysis; this data is shown below. 

```{r setup}
library(flipR)

simulation_data
```

As you can see, the orientation of rows and columns is not conducive for data manipulation,
wrangling, or plotting. This package provides tools for the user to solve this problem.


## flip()

Base R contains the function `t()`, which technically solves the problem explained above.
It switches the rows and columns of any matrix-like object inserted into the function.
However, it has the unlucky side effect of converting the data frame into a less useful
form; a matrix. In essence, `flip()` is a wrapper for base R `t()` but converts the result 
back into a more useful data frame. 

```{r flip}
#t() flips the rows and columns but results in a matrix
class(t(simulation_data))

#flip() returns a data frame
class(flip(simulation_data))
```

To aid the user in sliding this into their already-existing scripts, the `flip()` function,
as well as all other functions in this package, can optionally return the data in the form of a 
tibble from the [tibble package](https://tibble.tidyverse.org/) or data table from the [data.table package](https://cran.r-project.org/web/packages/data.table/index.html). 

```{r flip2}
#An additional argument can convert the output into other data frame forms
class(flip(simulation_data, "DT"))
class(flip(simulation_data, "TB"))
```

## The invert functions

In addition to flipping the rows and columns of a data frame object, one might need to
invert the order of rows or columns. For example, in the iris data set, it might be
more helpful to see the species and petal columns first. 

```{r invert_col}

head(iris)
head(invert_col(iris))
```

Alternatively, it might be helpful to see later rows first. For example, in the `USPersonalExpenditure`
data set one might want to see more Private Education first, so then I'd use the `invert_row()` function.

```{r invet_row}

head(USPersonalExpenditure)
head(invert_row(as.data.frame(USPersonalExpenditure)))
```

Just like the `flip()` function, the invert functions can also output data as tibbles or data tables.

```{r invert}
invert_col(iris, "DT")
invert_row(iris, "TB")
```

